---
title: "DATA607_Week2_Assignment"
author: "Don Padmaperuma"
date: "9/7/2019"
output: html_document
---
### Assignment – SQL and R
Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates,imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results(observations) and store them in a SQL database. Load the information into an R dataframe.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Load Appropriate Packages

```{r}
library(RMySQL)
```
### Establish Connection

```{r}

con <- dbConnect(MySQL(),
                  user = "root", 
                  password = "spring2019",
                  dbname = "movieratings",
                  host = "localhost"
                     )
```

Queries to get data from the SQL database
```{r}
query <- "select * from rating;"
res <- dbGetQuery(con, query)
res
```
### Load data into a R dataframe
```{r}

my_movie_ratings <- res
head(my_movie_ratings)

```
### Analyse Data using summary and graphs

```{r}
summary(my_movie_ratings)
```
```{r}
query1 <- "select * from rating;"
res <- dbGetQuery(con, query)
movie.rating <- data.frame(res)

query2 <- "select Name, The_Lion_King_2019 from rating;"

Lion_King_Rating <- dbGetQuery(con, query2)
Lion_King_Rating

```
```{r}
library(ggplot2)
ggplot(data = Lion_King_Rating, aes(Lion_King_Rating$Name, Lion_King_Rating$The_Lion_King_2019)) + geom_point(aes(color = "Rating 1-5")) + geom_smooth(method = "lm")
```
```{r}
query3 <- "select Name, Disney_Aladdin from rating;"
Disney_Aladdin_rating <- dbGetQuery(con, query3)
Disney_Aladdin_rating

```
```{r}
ggplot(data = Disney_Aladdin_rating, aes(Disney_Aladdin_rating$Name, Disney_Aladdin_rating$Disney_Aladdin))+ geom_point(aes(color = "Rating 1-5")) + geom_smooth(method = "lm")
```

### Disconnect from SQL database

```{r}
dbDisconnect(con)
```





