---
title: "DATA607_Week3_Assignment"
author: "Don Padmaperuma (Geeth)"
date: "9/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringr)
```

## Question 3  

```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
raw.data
```
```{r}
# Re arrange the vector using tools in Chapter 11 so that all the elements confirm to the standard first_name last_name
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```
```{r}
first_last <- unlist(sub("([[:alnum:]_]{1,}),  *([[:alnum:]_]{1,})", "\\2 \\1", sub("([[:alnum:]_]{1,})\\. "," ",name)))
first_last
```
```{r}
# Logical vector to indicate whether the character  has a title
title_vector <- c(str_detect(name, "(^[[:alnum:]_]{1,}\\.).*"))
name_title <- data.frame(name, title_vector)
name_title
```
```{r}
# Logical vector to indicate whether the character has a second name
second_name <- c(str_detect(name, " [[:alnum:]_]{1,}\\. "))
is_second <- data.frame(name, second_name)
is_second
```

## Question 4  

Discribe the type of strings and construct an example  
(1) **[0-9]+\$** 

**Any digit between 0 to 9 followed by $**  

```{r}
str_extract("ewqee3121$", "[0-9]+\\$")
```

(2) **\\b[a-z]{1,4}\\b**     

**set of strings with 1-4 lower case letters**  

```{r}
strings <- c("abcde", "abcd", "ab", "1234", "abc", "ABcd","12/03/2010", "10/27/1982", "text.txt", "data.csv", "essay.docx", "<header>My Web Site</header>")
unlist(str_extract_all(strings, "\\b[a-z]{1,4}\\b"))
```

(3) **.*?\\.txt$**  

**return strings ending with .txt**  

```{r}
unlist(str_extract_all(strings, ".*?\\.txt$" ))
```

(4) **\\d{2}/\\d{2}/\\d{4}**  

**string that contains digits in the format: xx/xx/xxxx**  

```{r}
unlist(str_extract_all(strings, "\\d{2}/\\d{2}/\\d{4}"))
```

(5) **<(.+?)>.+?</\\1>**  

**strings that contains opening and clossing tags as in html coding. Ex: <openTag>website</closeTag>**  
```{r}
unlist(str_extract_all(strings, "<(.+?)>.+?</\\1>"))
```
## Question 9  

```{r}
secret_message <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

uppercase <- paste(unlist(str_extract_all(secret_message, "[[A-Z].?]")), collapse = "")
lowercase <- paste(unlist(str_extract_all(secret_message, "[[a-z].?]")), collapse = "")
numbers <- paste(unlist(str_extract_all(secret_message, "[[0-9].?]")), collapse = "")

```

```{r}
uppercase
```
**Uppercase letters together makes a meaningful statement.**  

```{r}
lowercase
```
```{r}
numbers
```
**Decoded secret message**  
```{r}
decoded_message <- str_replace_all(uppercase, "[.]"," ")
decoded_message
```




      
