---
title: "DATA607_Week9"
author: "Don Padmaperuma (Geeth)"
date: "10/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment - Web APIs

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it into an R DataFrame.

## Libraries

```{r}
library(RCurl)
library(dplyr)
library(stringr)
library(httr)
library(jsonlite)
library(DT)
library(rvest)
library(kableExtra)
```


## Read in Json Data - Stephen King Books

Getting a list of best seller books where author is Stephen King from the Book API in New York Times  
```{r}
jsonFile1 <- "https://api.nytimes.com/svc/books/v3/reviews.json?author=Stephen+King&api-key=6Ll00ebXWXzaPjr4FQ3yMhevWwY0CHr0"

```

## Creating a Data frame

Creating a data frame using the json file that we just retrieved from NYT Books API
```{r}
Stephen_King_Books <- tbl_df(as.data.frame(fromJSON(jsonFile1)))
Stephen_King_Books %>% 
  kable ()%>%
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "500px", width = "100%")
```

## Read in Json Data - Best Seller History Books  

Getting a list of Best seller history books  
```{r}
jsonFile2 <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?&api-key=6Ll00ebXWXzaPjr4FQ3yMhevWwY0CHr0"
```

## Creating a Data frame  

Converting the json file to a dataframe
```{r}
BestSeller <- tbl_df(as.data.frame(fromJSON(jsonFile2)))
BestSeller %>% 
  kable ()%>%
  kable_styling(bootstrap_options = "striped", font_size = 10) %>% 
  scroll_box(height = "500px", width = "100%")
```

